---
title: "Kaggle Project"
author: "Tina Young, Anna Stone"
date: "`r Sys.Date()`"
output:
  word_document:
    toc: true
  html_document:
    toc: true
---

```{r setup, include=FALSE}
# This code chunk sets global options.

knitr::opts_chunk$set(echo = TRUE, message= F, warning = F)

# message= F and  warning = F suppress verbose information getting printed in 
# the compiled document.

# echo = T ensures that your code is visible.

# Load packages

library(tidyverse)
library(dplyr)

# Load data

train <- read.csv("C:/Users/young/OneDrive/Documents/Masters/IS 6489 Stats and Pred Analytics/Kaggle/train.csv")
test <- read.csv("C:/Users/young/OneDrive/Documents/Masters/IS 6489 Stats and Pred Analytics/Kaggle/test.csv")
```

# Introduction

The goal of this project is to create a model to predict house prices in Ames, Iowa. This training data consists of 1,460 observations and contains 79 explanatory variables. The predictive model will focus on 13 variables combined with a mixture of additive and interaction methods to predict the target variable, sale price. 

# Description of Data

```{r}
# Check for NAs
na <- colSums(is.na(train))
na[na > 0]
```

**Missing Data**

These values are legitimately missing and will be imputed with the median from the respective variable. 

- **LotFrontage**: 259 values
- **Electrical**: 1 value
- **MasVnrType and MasVnrArea**: 8 values each

**NA is Meaningful**

NA is being used to encode a missing feature. These values will be replaced with "none" instead of NA. 

- **Alley**: No alley access
- **BsmtQual, BsmtCond, BsmtExposure, BsmtFinType1, BsmtFinType2**: No Basement
- **FireplaceQu**: No Fireplace
- **GarageType, GarageYrBlt, GarageFinish, GarageQual, GarageCond**: No Garage
- **PoolQC**: No Pool
- **Fence**: No Fence
- **MiscFeature**: None

**Clean Data**

```{r}
#Replace NA values
## Define function to find the mode of a character vector
find_mode <- function(x){
  table(x) %>% 
    which.max() %>% 
    names()
}

train <- train |>
  mutate(GarageFinish = replace_na(data = GarageFinish, replace = "NoGarage"), 
         GarageQual = replace_na(data = GarageQual, replace = "NoGarage"), 
         GarageCond = replace_na(data = GarageCond, replace = "NoGarage"), 
         PoolQC = replace_na(data = PoolQC, replace = "NoPool"), 
         Fence = replace_na(data = Fence, replace = "NoFence"), 
         MiscFeature = replace_na(data = MiscFeature, replace = "None"), 
         GarageYrBlt = replace_na(data = GarageYrBlt, replace = 0), 
         Alley = replace_na(data = Alley, replace = "NoAlley"), 
         GarageType = replace_na(data = GarageType, replace = "NoGarage"), 
         BsmtQual = replace_na(data = BsmtQual, replace = "NoBasement"),
         BsmtCond = replace_na(data = BsmtCond, replace = "NoBasement"),
         BsmtExposure = replace_na(data = BsmtExposure, replace = "NoBasement"),
         BsmtFinType1 = replace_na(data = BsmtFinType1, replace = "NoBasement"),
         BsmtFinType2 = replace_na(data = BsmtFinType2, replace = "NoBasement"),
         MasVnrType = replace_na(data = MasVnrType, replace = find_mode(MasVnrType)), 
         FireplaceQu = replace_na(data = FireplaceQu, replace = "None"),
         Electrical = replace_na(data = Electrical, replace = find_mode(Electrical)),
         LotFrontage = replace_na(LotFrontage, median(LotFrontage, na.rm = T)), 
         MasVnrArea = replace_na(MasVnrArea, median(MasVnrArea, na.rm = T))) 

# Check for NAs
na <- colSums(is.na(train))
na[na > 0]
```

```{r message = F}
# Create scatterplot SalePrice ~ OverallQual to check if data is linear
ggplot(train, aes(as.numeric(OverallQual), SalePrice)) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  geom_smooth(se = F, col = 2) + # Local regression named LOESS
  labs(title = "SalePrice ~ OverallQual, with linear and local regression")
```

OverallQual should be factored. There is a non-linear relationship between OverallQual and SalesPrice. 

```{r}
# Factor OverallQual
train <- train |>
  mutate(OverallQual = factor(OverallQual))
```

# Selected Explanatory Variables
```{r results='hide'}
# Fit simple linear regression model

r1 <- lm(SalePrice ~ ExterQual, data = train)
summary(r1)

r2 <- lm(SalePrice ~ KitchenQual, data = train)
summary(r2)

r3 <- lm(SalePrice ~ OverallQual, data = train)
summary(r3)

r4 <- lm(SalePrice ~ GrLivArea, data = train)
summary(r4)

r5 <- lm(SalePrice ~ Neighborhood, data = train)
summary(r5)

r6 <- lm(SalePrice ~ BsmtQual, data = train)
summary(r6)

r7 <- lm(SalePrice ~ TotalBsmtSF, data = train)
summary(r7)

r8 <- lm(SalePrice ~ LotArea, data = train)
summary(r8)

r9 <- lm(SalePrice ~ OverallCond, data = train)
summary(r9)

r10 <- lm(SalePrice ~ YearBuilt, data = train)
summary(r10)

r11 <- lm(SalePrice ~ Fireplaces, data = train)
summary(r11)

r12 <- lm(SalePrice ~ LotShape, data = train)
summary(r12)

r13 <- lm(SalePrice ~ GarageCars, data = train)
summary(r13)

i <- lm(SalePrice ~ GarageCars * GrLivArea, data = train)
summary(i)
```

1) **ExterQual**: Evaluates the quality of the material on the exterior
2) **KitchenQual**: Kitchen quality
3) **OverallQual**: Rates the overall material and finish of the house
4) **GrLivArea**: Above grade (ground) living area square feet
5) **Neighborhood**: Physical locations within Ames city limits
6) **BsmtQual**: Height of the basement
7) **TotalBsmtSF**: Total square feet of basement area
8) **LotArea**: Lot size in square feet
9) **OverallCond**: Overall condition rating
10) **YearBuilt**: Original construction date
11) **Fireplaces**: Number of fireplaces
12) **LotShape**: General shape of property
13) **GarageCars**: Size of garage in car capacity

# Cross Validation

```{r}
# Randomly sample 70% of the rows
set.seed(123)
index <- sample(x = 1:nrow(train), size = nrow(train)*.7, replace = F)

head(index) # These are row numbers

# Subset train using the index to create train_fold
train_fold <- train[index, ]

# Subset the remaining row to create validation fold.
validation_fold <- train[-index, ]
```

```{r}
# Fit example model
model <- lm(SalePrice ~ GarageCars * GrLivArea + TotalBsmtSF + BsmtQual + ExterQual + KitchenQual + OverallQual + LotArea + Neighborhood + OverallCond + YearBuilt + Fireplaces + LotShape, data = train_fold)

summary(model)

# Get predictions for the train fold
predictions <- predict(model, newdata = train_fold)

# Create functions for calculating RMSE and R-squared
rmse <- function(observed, predicted) sqrt(mean((observed - predicted)^2))

R2 <- function(observed, predicted){
  TSS <- sum((observed - mean(observed))^2)
  RSS <- sum((observed - predicted)^2)
  1- RSS/TSS
}

# Calculate RMSE and R-squared for train fold
rmse(train_fold$SalePrice, predictions)
R2(train_fold$SalePrice, predictions)
```

An average error of $30,834.83 is observed in the training fold with the model explaining about 85% of the variance in sale price. This model meets the minimum requirements and can proceed to be utilized on the validation fold. 

```{r}
# Get predictions for the validation fold
predictions.v <- predict(model, newdata = validation_fold)

# Calculate RMSE and R-squared for validation fold
rmse(validation_fold$SalePrice, predictions.v)
R2(validation_fold$SalePrice, predictions.v)
```

An average error of $26,845.4 is observed in the training fold with the model explaining about 88% of the variance in sale price. This model performs better on the validation fold and is reliable for predicting sale prices on new data. 

# Predictions

```{r}
# Fit your model to the entire train set.
submission_model <- lm(SalePrice ~ GarageCars * GrLivArea + TotalBsmtSF + BsmtQual + ExterQual + KitchenQual + OverallQual + LotArea + Neighborhood + OverallCond + YearBuilt + Fireplaces + LotShape, data = train)

# Make exactly the same changes to the test set that you made to the train set.

test <- test |>
   mutate(OverallQual = factor(OverallQual), 
         GarageFinish = replace_na(data = GarageFinish, replace = "NoGarage"), 
         GarageQual = replace_na(data = GarageQual, replace = "NoGarage"), 
         GarageCond = replace_na(data = GarageCond, replace = "NoGarage"), 
         PoolQC = replace_na(data = PoolQC, replace = "NoPool"), 
         Fence = replace_na(data = Fence, replace = "NoFence"), 
         MiscFeature = replace_na(data = MiscFeature, replace = "None"), 
         GarageYrBlt = replace_na(data = GarageYrBlt, replace = 0), 
         Alley = replace_na(data = Alley, replace = "NoAlley"), 
         GarageType = replace_na(data = GarageType, replace = "NoGarage"), 
         BsmtQual = replace_na(data = BsmtQual, replace = "NoBasement"),
         BsmtCond = replace_na(data = BsmtCond, replace = "NoBasement"),
         BsmtExposure = replace_na(data = BsmtExposure, replace = "NoBasement"),
         BsmtFinType1 = replace_na(data = BsmtFinType1, replace = "NoBasement"),
         BsmtFinType2 = replace_na(data = BsmtFinType2, replace = "NoBasement"),
         MasVnrType = replace_na(data = MasVnrType, replace = find_mode(MasVnrType)), 
         FireplaceQu = replace_na(data = FireplaceQu, replace = "None"),
         Electrical = replace_na(data = Electrical, replace = find_mode(Electrical)), 
         LotFrontage = replace_na(LotFrontage, median(LotFrontage, na.rm = T)), 
         MasVnrArea = replace_na(MasVnrArea, median(MasVnrArea, na.rm = T)), 
         TotalBsmtSF = replace_na(data = TotalBsmtSF, replace = median(TotalBsmtSF, na.rm = TRUE)),
         KitchenQual = replace_na(data = KitchenQual, replace = find_mode(KitchenQual)),
         GarageCars = replace_na(data = GarageCars, replace = median(GarageCars, na.rm = TRUE)))

# Check there are no missing observations for your selected predictors in the test set.
test |>
  select(GarageCars, GrLivArea, TotalBsmtSF, BsmtQual, ExterQual, KitchenQual, OverallQual, LotArea, Neighborhood, OverallCond, YearBuilt, Fireplaces, LotShape) |> 
  summarize_all(~sum(is.na(.)))

# Make predictions for the test set.
submission_predictions <- predict(submission_model, newdata = test) # Use the newdata argument!

head(submission_predictions)

# Format your submission file.

submission <- test |>
  select(Id) |>
  mutate(SalePrice = submission_predictions)

# Check
head(submission)

# write to csv
write.csv(submission, "kaggle_group_submission.csv", row.names = F)
```

# Kaggle Score and Rank

-**Score**: 0.14390


-**Rank**: 2883

# List of Contributions

-**Anna**: Editor, provided code for models, commented on model fit, formatted and submitted to Kaggle

-**Tina**: 

